---
output: github_document
---

#### *Marcus W. Beck, marcusb@sccwrp.org, Robert Butler, r7butler@yahoo.com, Raphael D. Mazor, raphaelm@sccwrp.org, Mark Engeln*

# PHABMetrics

[![Travis-CI Build Status](https://travis-ci.org/SCCWRP/PHABMetrics.svg?branch=master)](https://travis-ci.org/SCCWRP/PHABMetrics)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/SCCWRP/PHABMetrics?branch=master&svg=true)](https://ci.appveyor.com/project/SCCWRP/PHABMetrics)

This package provides functions to calculate PHAB metrics using field data.

# Installing the package

The development version of this package can be installed from Github:

```{r eval = F, message = F}
install.packages('devtools')
library(devtools)
install_github('SCCWRP/PHABMetrics')
library(PHABMetrics)
```

# Usage

```{r, message = F, warning = F, echo = F}
devtools::load_all()
```

Input data format:

```{r}
head(sampdat)
```

The core  function is `phabmetrics()`.  Calulcate PHAB metrics with sample data:

```{r}
alldat <- phabmetrics(sampdat)
```

```{r, echo = F}
mets <- names(alldat)
mets <- gsub('\\..*$', '', mets)
mets <- unique(mets)[!unique(mets) %in% 'StationCode']
```

The following `r length(mets)` metrics are calculated:
```{r, echo = F}
mets
```

```{r}
alldat
```

# Required data checks

* For every function, make sure there are no duplicate or conflicting values for every unique combination of `id`, `LocationCode`, `AnalyteName`, and `VariableResult` (or `Result`).  This should be specific to the metric classes just to be safe.  For example, every combination should have only one entry in `VariableResult` for `AnalyteName %in% c('Microalgae Thickness', 'Macrophyte Cover', 'Macroalgae Cover, Attached', 'Macroalgae Cover, Unattached')` for the algae metrics.  The `algae.R` function will remove duplicate entries but a checker should be built that verifies a unique value can be determined.  

* Required column names, see those in `sampdat`.  

* Check for required values in `AnalyteName` (note that `chkinp()` can check of the columns exist but we'll need a checker on data input to check for these and only these): 
     * `c('Microalgae Thickness', 'Macrophyte Cover', 'Macroalgae Cover, Attached', 'Macroalgae Cover, Unattached')` for `algae()`
     * `c('Bankfull Height', 'Bankfull Width', 'StationWaterDepth', 'Wetted Width')` for `bankmorph()`
     * `c('Cascade/Falls', 'Dry', 'Glide', 'Pool', 'Rapid', 'Riffle', 'Run'))` for `channelmorph()` 
     * `c(Slope', 'Length, Segment', 'Elevation Difference', 'Bearing', 'Proportion', 'Length, Reach')` for `channelsinuosity()`
     * `c('Canopy Cover')` for `densiometer()`
     * `c('Distance from Bank', 'StationWaterDepth', 'Velocity', 'Distance, Float', 'Float Time', 'Wetted Width')` for `flow()`
     * `c('Fish Cover Macrophytes', 'Fish Cover Artificial Structures', 'Fish Cover Boulders', 'Fish Cover Filamentous Algae', 'Fish Cover Woody Debris >0.3 m', 'Fish Cover Live Trees/Roots', 'Fish Cover Overhang.Veg', 'Fish Cover Woody Debris <0.3 m', 'Fish Cover Undercut Banks')` for `habitat()`
     * `c('Riparian Bridges/Abutments', 'Riparian Buildings', 'Riparian Landfill/Trash', 'Riparian Logging', 'Riparian Mining', 'Riparian Orchards/Vineyards', 'Riparian Park/Lawn', 'Riparian Pasture/Range', 'Riparian Pavement', 'Riparian Pipes', 'Riparian Road', 'Riparian Row Crops', 'Riparian Vegetation Management', 'Riparian Wall/Dike')` for `disturbance()`
     * `c('Riffle/Run Channel Alteration', 'Riffle/Run Epifaunal Substrate', 'Riffle/Run Sediment Deposition', 'Dominant Land Use', 'Evidence of Fire', 'Evidence of Recent Rainfall')` for `misc()`
     * `c('Bank Stability')` for `bankstability()`
     * `c("Alkalinity as CaCO3", "Oxygen, Dissolved", "pH", "Salinity", "SpecificConductivity", "Temperature", "Turbidity")` for `quality()`
     * `c('Riparian GroundCover Barren', 'Riparian GroundCover NonWoody Plants', 'Riparian GroundCover Woody Shrubs', 'Riparian Lower Canopy All Vegetation', 'Riparian Upper Canopy All Trees', 'Riparian Lower Canopy All Vegetation', 'Riparian Upper Canopy All Trees', 'Riparian GroundCover Woody Shrubs', 'Riparian GroundCover NonWoody Plants')` for `ripveg()` 
     * `c('Substrate Size Class', 'Embeddedness', 'CPOM')` for `substrate()`
     
* Check for required values in `LocationCode` (note that `chkinp()` can check if the columns exist but we'll need a checker on data input to check for these and only these)

* Maybe we need to add a checker to make sure all values in each field are present but with appropriate NA values for `Result`, `VariableResult`, this can be done with `tidyr::complete()` but may be unnecessary since this will increase data volume 